Move Schedules: Fast persistence computations in coarse dynamic settings\thanks{\textbf{Funding:} This work was partially supported by the National Science Foundation through grants CCF-2006661 and CAREER award DMS-1943758.}

### 2021-09-28 move_schedules.tex
Last updates before overhaul of last part of section 3

  		w \kern\rowsp & \; 1 &  1 &    \; \botstrut \\

@@

@@

Notice that homological critical events can only occur when the filtration order changes, though not all filtration changes result in homological critical events.
It follows that detecting these homological events reduces to computing matrix decompositions at a finite set of time points interleaved between filtration changes. 
At the finest scale, changes to the filtration order manifest as adjacent transpositions. 

@@

  		 & \; 1 & \textbf{1} & \; \botstrut \\

@@

  		& \; 1 & \textbf{1} &  \; \botstrut \\

@@

@@

\end{algorithm}
If one has a pair of filtrations $(K_0, K_1)$ each with $m$ simplices, and we assume that each simplex $\sigma_i$ switches its relative ordering with another simplex $\sigma_j$ at most once in some ordered schedule of transpositions, then the number of intermediate persistence diagrams is bounded above by $O(m^2)$ using transpositions, whereas with move operations this bound  is at most $O(m)$.  
\\
\\
\noindent \textbf{Example:} We continue the example used in sections~\ref{sec:vineyards} and~\ref{sec:reduction} to illustrate moves. Consider moving the edge $a$ to the position of edge $c$ in the filtration. The donor columns introduced in~\cite{busaryev2010tracking} are shown on the left side of each matrix. Note that using vineyards, the equivalent permutation to the decomposition would require $4$ column operations on both $R_1$ and $V_1$, respectively, whereas a single move operation accomplishes this permutation using only $2$ column operations per matrix. 
\begin{displaymath}
	\def\rowsp{0.4em}
	\def\arraystretch{0.95}
	\makeatletter\setlength\BA@colsep{1.2pt}\makeatother
	%
	\begin{blockarray}{cc}{d_R}
	\kern-0.2em & a \\
		\begin{block}{c[c]}
  		u \kern \rowsp & \; 1 \; \topstrut \\
  		v \kern \rowsp & \;  \; \\
  		w \kern \rowsp & \; 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \;
	\begin{blockarray}{cccc}{R_1}
	 \kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		u \kern \rowsp & \; 1 & 1 &  \; \topstrut \\
  		v \kern \rowsp & \;  & 1 &  \; \\
  		w \kern \rowsp & \; 1 &  &  \; \botstrut \\
		\end{block}
	\end{blockarray}
	%
\; \to
	%
\begin{blockarray}{cc}{}
	\kern-0.2em & b \\
		\begin{block}{c[c]}
  		  & \; 1 \; \topstrut \\
  		 & \;  1 \; \\
  		 & \; \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \;
\begin{blockarray}{cccc}{}
	\kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		  & \; 1 &    &  \; \topstrut \\
  		  & \;    & 1 &  \; \\
  		 & \; 1 & 1 &  \; \botstrut \\
		\end{block}
	\end{blockarray}
	%
\; \to
	%
	\begin{blockarray}{cc}{}
	\kern-0.2em & c \\
		\begin{block}{c[c]}
  		  & \; \hphantom{1} \; \topstrut \\
  		 & \;  \hphantom{1} \; \\
  		 & \; \hphantom{1} \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \; 
	\begin{blockarray}{cccc}{}
	\kern-0.2em & a & b &  c \\
		\begin{block}{c[ccc]}
  		  & \; 1   &    &   1  \; \topstrut \\
  		 & \;       &  1  &  1  \; \\
  		 & \; 1   &   1   &     \; \botstrut \\
		\end{block}
	\end{blockarray}
%
\; \xrightarrow{P} 
\begin{blockarray}{cc}{}
	\kern-0.2em & c \\
		\begin{block}{c[c]}
  		  & \; \hphantom{1} \; \topstrut \\
  		 & \; \hphantom{1} \; \\
  		 & \; \hphantom{1} \; \botstrut \\
		\end{block}
	\end{blockarray}
\; \;
\begin{blockarray}{cccc}{}
	\kern-0.2em & b & c & a  \\
		\begin{block}{c[ccc]}
  		  & \;  & 1 &  1\; \topstrut \\
  		 & \; 1 & 1 &  \; \\
  		 & \; 1 &  & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
%
\; \xrightarrow{d_R} 
\begin{blockarray}{cccc}{R_1}
\kern-0.2em & b & c & a  \\
	\begin{block}{c[ccc]}
		u \kern \rowsp & \;  & 1 &  \; \topstrut \\
		v \kern \rowsp & \; 1 & 1 &  \; \\
		w \kern \rowsp & \; 1 &  &  \; \botstrut \\
	\end{block}
	% \undermat{B}{ & & }  
\end{blockarray}
\end{displaymath}

@@

% Start V example 
\begin{displaymath}
	\def\rowsp{0.4em}
	\def\arraystretch{0.95}
	\makeatletter\setlength\BA@colsep{1.2pt}\makeatother
	%
	\begin{blockarray}{cc}{d_V}
	\kern-0.2em & a \\
		\begin{block}{c[c]}
  		 & \; 1 \; \topstrut \\
  		& \;  \; \\
  		& \;  \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \;
	\begin{blockarray}{cccc}{V_1}
	 \kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		 & \; 1 & 1 &  1\; \topstrut \\
  		& \;  & 1 &  1\; \\
  		& \; &  & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	%
\; \to
	%
\begin{blockarray}{cc}{}
	\kern-0.2em & b \\
		\begin{block}{c[c]}
  		 & \; 1 \; \topstrut \\
  		& \;  1 \; \\
  		& \; \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \;
\begin{blockarray}{cccc}{}
	 \kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		 & \; 1 & &  1\; \topstrut \\
  		& \;  & 1 &  1\; \\
  		& \; &   & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	%
\; \to
	%
	\begin{blockarray}{cc}{}
	\kern-0.2em & c \\
		\begin{block}{c[c]}
  		 & \; 1 \; \topstrut \\
  		& \;  1 \; \\
  		& \; 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \;
	\begin{blockarray}{cccc}{}
	\kern-0.2em & a & b &  c \\
		\begin{block}{c[ccc]}
  		 & \; 1 & &   \; \topstrut \\
  		& \;  & 1 &   \; \\
  		& \;  &   & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
%
\; \xrightarrow{P} \\
% 
\begin{blockarray}{cc}{}
	\kern-0.2em & c \\
		\begin{block}{c[c]}
  		 & \; 1 \; \topstrut \\
  		& \;  1 \; \\
  		& \; 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
\; \;
\begin{blockarray}{cccc}{}
	\kern-0.2em & b & c & a  \\
		\begin{block}{c[ccc]}
  		 & \; 1 & &\; \topstrut \\
  		& \;  & 1 &  \; \\
  		& \; &  & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
%
\; \xrightarrow{d_V}
% 
	\begin{blockarray}{cccc}{V_1}
	\kern-0.2em & b & c & a  \\
		\begin{block}{c[ccc]}
  		b \kern \rowsp & \; 1 & & 1 \; \topstrut \\
  		c \kern \rowsp & \; & 1 &  1 \; \\
  		a \kern \rowsp & \; &  & 1  \; \botstrut \\
		\end{block}
	\end{blockarray}
\end{displaymath}
\\
\noindent
We recall an important claim given in~\cite{busaryev2010tracking} on the effect that move operations have on the status of simplices in the pairing. Recall from section~\ref{sec:background} that simplices which create new homology classes are called \emph{creators} and simplices that destroy homology classes are called \emph{destroyers}.
%\begin{proposition}[Locality of moves~\cite{busaryev2010tracking}]
%	Other than the simplex $\sigma_i$ being moved, $\mathrm{MoveRight}(i,j)$ cannot change a destroyer to a creator, $\mathrm{MoveLeft}(i,j)$ cannot change a creator to destroyer, and neither move operation changes the status of any simplex in $K$ outside of the interval $[i,j]$.  
%\end{proposition}
\noindent The effect of the movement on intermediate simplices depends on the direction of the movement. If $i < j$ (respectively, $j < i$), all simplices at positions $k \in [i+1\;:\;j] $ are shifted down (respectively, up) by $1$. 
The pseudo-code for the \emph{MoveRight} operation is given in Algorithm~\ref{alg:mr}, and the pseudo-code for the \emph{MoveLeft} is given in Algorithm~\ref{alg:ml}.

@@

	Now consider using the Move($i$,$j$) in Algorithm~\ref{alg:mr} for a single move operation, as a replacement for the sequence of transpositions above. Here again, the dominant cost again are the column operations (line 5), since although permutations take $O(\lvert i - j\rvert)$ time we require only $1$ such permutation per movement. 
	If we allocate an extra $O(m)$ memory space during the movement, notice that lines (7) and (7) in both Restore* and Move* can be completed $O(1)$ via a pointer swapping argument. 
	Since a move operation requires at most $1$ column operation for each index in $[i: j]$, moving a column from $i$ to $j$ where $i < j$ requires at most $2(\lvert i - j \rvert)$ column operations. The claimed inequality follows. 
	%Identifying non-zero entries in the matrix occurs once, which takes $O(\lvert i - j \rvert \log(m))$ time.
\end{proof}
In terms of bounds, we  have $\lvert \widetilde{S}_F \rvert \leq \lvert S_F \rvert$, and the associated coarsened schedule $\widetilde{S}_F$ can be computed in $O(m)$ time. 
However, the coarsening depends entirely on the choice of $F$ and the upper bound from~\ref{eq:sf_m2} remains, as its always possible that there are no contiguous subsequences to collapse. If a straight line homotopy is used, the number of collapsible subsequences may depend on the simplices categorization as critical points; transpositions between pairs of simplices which trigger  changes in homology tend to be more expensive. Cohen-Steiner~\cite{cohen2006vines} referred to these as \emph{switches}, see~\cite{edelsbrunner2000topological} for more details.

@@

@@

 \begin{algorithm}[h]
	\caption{Sorting algorithm}\label{alg:sorting}
	\setstretch{1.05}
    \begin{algorithmic}[1]
    	% \Require Fixed $p' \in S_m$, $\mathcal{L} = \mathrm{LIS}(p')$ 
    	\Function{LCS-Sort}{$p$, $q$}
    		\State $\bar{p} \gets q' \circ p$
    		\State $\mathcal{L} \gets \mathrm{LCS}(p,q) = \mathrm{LIS}(\bar{p})$ \Comment{$O(\log \log m)$}
    		\State $\mathcal{D}, \mathcal{S} \gets \bar{p} \setminus \mathcal{L}, \emptyset$
    		\While{$\mathcal{D}$\text{ is not empty }}
    		    \State $\sigma \gets \text{arbitrary element in }\mathcal{D}$
    		    %\State $\tau, \rho \gets \mathcal{T}_p(\sigma), \mathcal{T}_n(\sigma)$
    		    %\State $i, i_p, i_n \gets \mathcal{I}_p(\sigma), \, \mathcal{I}_p(T_{\mathrm{pred}}(\sigma)), \, \mathcal{I}_p(T_{\mathrm{succ}}(\sigma))$
    		    \State $i, i_p, i_n \gets \bar{p}^{-1}(\sigma), \bar{p}^{-1}(T_{\mathrm{pred}}(\sigma)), \bar{p}^{-1}(T_{\mathrm{succ}}(\sigma))$%\Comment{$O(\log \log m)$}
%    		    \If{ $p_{\sigma} < p_{\tau}$ }
%	    		    \State $j \gets \text{arbitrary element in } [\, p_{\tau}, \, p_{\rho})$
%    		    \Else{ $p_{\rho} < p_{\sigma}$}
%    		    	\State $j \gets \text{arbitrary element in } (\, p_{\tau}, \, p_{\rho}]$
%    		    \EndIf
    		    \If{ $i < i_p$ } 
    		    	\State $j \gets \text{arbitrary element in } [\, i_p, \, i_n)$
    		    \Else{ $i_n < i$ }
	    		    \State $j \gets \text{arbitrary element in } (\, i_p, \, i_n]$ 
    		    \EndIf
    		    \State $\mathcal{S}, \mathcal{D}, \mathcal{T} \gets \mathcal{S} \cup (i, \, j), \mathcal{D} \setminus \sigma, \mathcal{T} \cup \sigma$%\Comment{$O(\log \log m)$}
    		    % \State $p, p^{-1} \gets m_{ij} \circ p, p^{-1} \circ m_{ij}^{-1}$
    		    \State $\bar{p}^{-1} \gets \bar{p}^{-1} \circ m_{ij}^{-1}$ %\Comment{$O(m)$}
    		 \EndWhile
    		 \State \Return{$\mathcal{S}$}
    	\EndFunction
	\end{algorithmic}
\end{algorithm}

>>>>

  		w \kern\rowsp & \; 1 &  \fbox{1} &    \; \botstrut \\

@@

  		v \kern\rowsp & \;     & 1 & \fbox{1} \; \\

@@

Notice that homological critical events can only occur when the filtration order changes, though the converse isn't true.
Thus, detecting these all such events reduces to computing matrix decompositions at a finite set of time points interleaved between changes in the filtration order. 
At the finest scale, these changes manifest as adjacent transpositions. 

@@

  		 & \; 1 & 1 & \; \botstrut \\

@@

  		& \; 1 & \fbox{1} &  \; \botstrut \\

@@

  		  & \; 1 & 1 & 1 \; \topstrut \\

@@

\end{algorithm}
If one has a pair of filtrations $(K_0, K_1)$ each with $m$ simplices, and we assume that each simplex $\sigma_i$ switches its relative ordering with another simplex $\sigma_j$ at most once in some ordered schedule of transpositions, then the number of intermediate persistence diagrams is bounded above by $O(m^2)$ using transpositions, whereas with move operations this bound  is at most $O(m)$.  
\\
\\
\noindent \textbf{Example:} We continue the example used in sections~\ref{sec:vineyards} and~\ref{sec:reduction} to illustrate moves. Consider moving the edge $a$ to the position of edge $c$ in the filtration. The donor columns introduced in~\cite{busaryev2010tracking} are shown on the left side of each matrix. Note that using vineyards, the equivalent permutation to the decomposition would require $4$ column operations on both $R_1$ and $V_1$, respectively, whereas a single move operation accomplishes this permutation using only $2$ column operations per matrix. 
\begin{displaymath}
	\def\rowsp{0.4em}
	\def\arraystretch{0.95}
	\makeatletter\setlength\BA@colsep{1.2pt}\makeatother
	%
	\begin{blockarray}{cc}{d_R}
	\kern-0.2em & a \\
		\begin{block}{c[c]}
  		u \kern \rowsp & \; 1 \; \topstrut \\
  		v \kern \rowsp & \;  \; \\
  		w \kern \rowsp & \; 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \;
	\begin{blockarray}{cccc}{R_1}
	 \kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		u \kern \rowsp & \; 1 & 1 &  \; \topstrut \\
  		v \kern \rowsp & \;  & 1 &  \; \\
  		w \kern \rowsp & \; 1 &  &  \; \botstrut \\
		\end{block}
	\end{blockarray}
	%
\; \to
	%
\begin{blockarray}{cc}{}
	\kern-0.2em & b \\
		\begin{block}{c[c]}
  		  & \; 1 \; \topstrut \\
  		 & \;  1 \; \\
  		 & \; \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \;
\begin{blockarray}{cccc}{}
	\kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		  & \; 1 &    &  \; \topstrut \\
  		  & \;    & 1 &  \; \\
  		 & \; 1 & \fbox{1} &  \; \botstrut \\
		\end{block}
	\end{blockarray}
	%
\; \to
	%
	\begin{blockarray}{cc}{}
	\kern-0.2em & c \\
		\begin{block}{c[c]}
  		  & \; \hphantom{1} \; \topstrut \\
  		 & \;  \hphantom{1} \; \\
  		 & \; \hphantom{1} \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \; 
	\begin{blockarray}{cccc}{}
	\kern-0.2em & a & b &  c \\
		\begin{block}{c[ccc]}
  		  & \; 1   &    &   1  \; \topstrut \\
  		 & \;       &  1  &  1  \; \\
  		 & \; 1   &   \fbox{1}   &     \; \botstrut \\
		\end{block}
	\end{blockarray}
%
\; \xrightarrow{P} 
\begin{blockarray}{cc}{}
	\kern-0.2em & c \\
		\begin{block}{c[c]}
  		  & \; \hphantom{1} \; \topstrut \\
  		 & \; \hphantom{1} \; \\
  		 & \; \hphantom{1} \; \botstrut \\
		\end{block}
	\end{blockarray}
\; \;
\begin{blockarray}{cccc}{}
	\kern-0.2em & b & c & a  \\
		\begin{block}{c[ccc]}
  		  & \;  & 1 &  1\; \topstrut \\
  		 & \; 1 & 1 &  \; \\
  		 & \; 1 &  & \fbox{1} \; \botstrut \\
		\end{block}
	\end{blockarray}
%
\; \xrightarrow{d_R} 
\begin{blockarray}{cccc}{R_1}
\kern-0.2em & b & c & a  \\
	\begin{block}{c[ccc]}
		u \kern \rowsp & \;  & 1 &  \; \topstrut \\
		v \kern \rowsp & \; 1 & 1 &  \; \\
		w \kern \rowsp & \; 1 &  &  \; \botstrut \\
	\end{block}
	% \undermat{B}{ & & }  
\end{blockarray}
\end{displaymath}

@@

% Start V example 
\begin{displaymath}
	\def\rowsp{0.4em}
	\def\arraystretch{0.95}
	\makeatletter\setlength\BA@colsep{1.2pt}\makeatother
	%
	\begin{blockarray}{cc}{d_V}
	\kern-0.2em & a \\
		\begin{block}{c[c]}
  		 & \; 1 \; \topstrut \\
  		& \;  \; \\
  		& \;  \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \;
	\begin{blockarray}{cccc}{V_1}
	 \kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		 & \; 1 & \fbox{1} &  \fbox{1}\; \topstrut \\
  		& \;  & 1 &  1\; \\
  		& \; &  & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	%
\; \to
	%
\begin{blockarray}{cc}{}
	\kern-0.2em & b \\
		\begin{block}{c[c]}
  		 & \; 1 \; \topstrut \\
  		& \;  1 \; \\
  		& \; \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \;
\begin{blockarray}{cccc}{}
	 \kern-0.2em & a & b & c  \\
		\begin{block}{c[ccc]}
  		 & \; 1 & &  \fbox{1}\; \topstrut \\
  		& \;  & 1 &  1\; \\
  		& \; &   & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	%
\; \to
	%
	\begin{blockarray}{cc}{}
	\kern-0.2em & c \\
		\begin{block}{c[c]}
  		 & \; 1 \; \topstrut \\
  		& \;  1 \; \\
  		& \; 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
	\; \;
	\begin{blockarray}{cccc}{}
	\kern-0.2em & a & b &  c \\
		\begin{block}{c[ccc]}
  		 & \; 1 & &   \; \topstrut \\
  		& \;  & 1 &   \; \\
  		& \;  &   & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
%
\; \xrightarrow{P} \\
% 
\begin{blockarray}{cc}{}
	\kern-0.2em & a \\
		\begin{block}{c[c]}
  		 & \; 1 \; \topstrut \\
  		& \;  1 \; \\
  		& \; 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
\; \;
\begin{blockarray}{cccc}{}
	\kern-0.2em & b & c & a  \\
		\begin{block}{c[ccc]}
  		 & \; 1 & &\; \topstrut \\
  		& \;  & 1 &  \; \\
  		& \; &  & 1 \; \botstrut \\
		\end{block}
	\end{blockarray}
%
\; \xrightarrow{d_V}
% 
	\begin{blockarray}{cccc}{V_1}
	\kern-0.2em & b & c & a  \\
		\begin{block}{c[ccc]}
  		b \kern \rowsp & \; 1 & & 1 \; \topstrut \\
  		c \kern \rowsp & \; & 1 &  1 \; \\
  		a \kern \rowsp & \; &  & 1  \; \botstrut \\
		\end{block}
	\end{blockarray}
\end{displaymath}
\\
\noindent
We recall an important claim given in~\cite{busaryev2010tracking} on the effect that move operations have on the status of simplices in the pairing. Recall from section~\ref{sec:background} that simplices which create new homology classes are called \emph{creators} and simplices that destroy homology classes are called \emph{destroyers}.
%\begin{proposition}[Locality of moves~\cite{busaryev2010tracking}]
%	Other than the simplex $\sigma_i$ being moved, $\mathrm{MoveRight}(i,j)$ cannot change a destroyer to a creator, $\mathrm{MoveLeft}(i,j)$ cannot change a creator to destroyer, and neither move operation changes the status of any simplex in $K$ outside of the interval $[i,j]$.  
%\end{proposition}
\noindent The effect of the movement on intermediate simplices depends on the direction of the movement. If $i < j$ (respectively, $j < i$), all simplices at positions $k \in [i+1\;:\;j] $ are shifted down (respectively, up) by $1$. 
The pseudo-code for the \emph{MoveRight} operation is given in Algorithm~\ref{alg:mr}, and the pseudo-code for the \emph{MoveLeft} is given in Algorithm~\ref{alg:ml}.

@@

	Now consider using the MoveRight($i$,$j$) for a single move operation, as a replacement for the sequence of transpositions above. Here again, the dominant cost again are the column operations (line 5), since although permutations take $O(\lvert i - j\rvert)$ time we require only $1$ such permutation per movement. 
	If we allocate an extra $O(m)$ memory space during the movement, notice that lines (4), and (7) of Restore* (but not line (2)) and line (7) of Move* in Algorithm~\ref{alg:mr} can be completed $O(1)$ via a pointer swapping argument. 
	Since a move operation requires at most $1$ column operation for each index in $[i: j]$, moving a column from $i$ to $j$ where $i < j$ requires at most $2(\lvert i - j \rvert)$ column operations. The claimed inequality follows. 
	%Identifying non-zero entries in the matrix occurs once, which takes $O(\lvert i - j \rvert \log(m))$ time.
\end{proof}
In terms of bounds, we have $\lvert \widetilde{S}_F \rvert \leq \lvert S_F \rvert$, and the associated coarsened schedule $\widetilde{S}_F$ can be computed in $O(m)$ time. 
However, the coarsening depends entirely on the choice of $F$ and the upper bound from~\ref{eq:sf_m2} remains, as its always possible that there are no contiguous subsequences to collapse. If a straight line homotopy is used, the number of collapsible subsequences may depend on the simplices categorization as critical points; transpositions between pairs of simplices which trigger  changes in homology tend to be more expensive. Cohen-Steiner~\cite{cohen2006vines} referred to these as \emph{switches}, see~\cite{edelsbrunner2000topological} for more details.

@@

 \algnewcommand{\IfThenElse}[3]{% \IfThenElse{<if>}{<then>}{<else>}
  \State \algorithmicif\ #1\ \algorithmicthen\ #2\ \algorithmicelse\ #3}

  \begin{algorithm}[h]
	\caption{Sorting algorithm}\label{alg:sorting}
	\setstretch{1.05}
    \begin{algorithmic}[1]
    	% \Require Fixed $p' \in S_m$, $\mathcal{L} = \mathrm{LIS}(p')$ 
    	\Function{LCS-Sort}{$p$, $q$}
    		\State $\bar{p} \gets q' \circ p$
    		\State $\mathcal{L} \gets \mathrm{LCS}(p,q) = \mathrm{LIS}(\bar{p})$ \Comment{$O(\log \log m)$}
    		\State $(\, \mathcal{S}, \, \mathcal{D}, \, \mathcal{T} \, )\gets (\, \emptyset, \, \bar{p} \setminus \mathcal{L}, \, \mathcal{L} \,)$
    		\While{$\mathcal{D}$\text{ is not empty }}
    		    \State $\sigma \gets \text{arbitrary element in }\mathcal{D}$
    		    %\State $\tau, \rho \gets \mathcal{T}_p(\sigma), \mathcal{T}_n(\sigma)$
    		    %\State $i, i_p, i_n \gets \mathcal{I}_p(\sigma), \, \mathcal{I}_p(T_{\mathrm{pred}}(\sigma)), \, \mathcal{I}_p(T_{\mathrm{succ}}(\sigma))$
    		    \State $(\,i, \,i_p, \, i_n \,) \gets ( \, \bar{p}^{-1}(\sigma), \, \bar{p}^{-1}(T_{\mathrm{pred}}(\sigma)), \, \bar{p}^{-1}(T_{\mathrm{succ}}(\sigma)) \, )$%\Comment{$O(\log \log m)$}
%    		    \If{ $p_{\sigma} < p_{\tau}$ }
%	    		    \State $j \gets \text{arbitrary element in } [\, p_{\tau}, \, p_{\rho})$
%    		    \Else{ $p_{\rho} < p_{\sigma}$}
%    		    	\State $j \gets \text{arbitrary element in } (\, p_{\tau}, \, p_{\rho}]$
%    		    \EndIf
				% \IfThenElse{$i < i_p$}{$[\, i_p, \, i_n)$}{$(\, i_p, \, i_n]$}
				% $\text{Pick } j \in [\, i_p, \, i_n)$ \If{ $i < i_p$ } \Else $(\, i_p, \, i_n]$ \EndIf
    		    \If{ $i < i_p$ } 
    		    	\State $j \gets \text{arbitrary element in } [\, i_p, \, i_n)$
    		    \Else{ $i_n < i$ }
	    		    \State $j \gets \text{arbitrary element in } (\, i_p, \, i_n]$ 
    		    \EndIf
    		    \State $(\, \mathcal{S}, \mathcal{D}, \mathcal{T} \,) \gets (\, \mathcal{S} \cup (i, \, j), \, \mathcal{D} \setminus \sigma, \, \mathcal{T} \cup \sigma \, )$%\Comment{$O(\log \log m)$}
    		    % \State $p, p^{-1} \gets m_{ij} \circ p, p^{-1} \circ m_{ij}^{-1}$
    		    \State $\bar{p}^{-1} \gets \bar{p}^{-1} \circ m_{ij}^{-1}$ %\Comment{$O(m)$}
    		 \EndWhile
    		 \State \Return{$\mathcal{S}$}
    	\EndFunction
	\end{algorithmic}
\end{algorithm}

@@
